Assignment # 3
========================================================
Qustion#1
Download the file
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/rprog%2Fdata%2FProgAssignment3-data.zip"

if (!file.exists("Assignment3-data.zip")) {
    download.file(fileUrl, destfile = "Assignment3-data.zip", method = "wget")
}
```

Unzip the file

```{r}
if (!file.exists("Assignment3-data")) {
    unzip("Assignment3-data.zip", overwrite = TRUE)
}
```


```{r}
getwd()
```


1 Plot the 30-day mortality rates for heart attack
```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
```

Number of columns 
```{r}
ncol(outcome)
```


```{r}
names(outcome)
```

To make a simple histogram of the 30-day death rates from heart attack (column 11 in the outcome dataset), run
```{r}
names(outcome)[11]
outcome[, 11] <- as.numeric(outcome[, 11])
## You may get a warning about NAs being introduced; that is okay
```

```{r fig.width=7, fig.height=6}
hist(outcome[, 11])
```


##2 Finding the best hospital in a state

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can be one of “heart attack”, “heart failure”, or “pneumonia”. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.


```{r}
best  <- function(state, outcome) {
  ## Read outcome data 
  df  <- read.csv('./outcome-of-care-measures.csv',
      colClasses=c('Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia'='character',
                   'Hospital.Name'='character'))
  ## Check that state and outome are valid
  outcomes   <- c("heart attack","heart failure","pneumonia")
  if (!(state %in% df$State)) {
    stop("invalid state")
  }
  
  if (!(outcome %in% outcomes)) {
    stop("invalid outcome")
  }
  
  ## Return hospital name in that state with lowest 30-day death 
  ## rate
  df1  <- df[df$State == state,]
  if (outcome == "heart attack") {
    return(min(df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T), 2]))
  }
  
  if (outcome == "heart failure") {
    return(min(df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure,na.rm = T), 2]))
  }

  if (outcome == "pneumonia") {
    return(min(df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia,na.rm = T), 2]))
  }
  
}
```

Unit tests

```{r}
best("TX", "heart attack")
best("TX", "heart failure")
best("MD", "heart attack")
best("MD", "pneumonia")
best("BB", "heart attack")
best("NY", "hert attack")
best("SC", "heart attack")
best("NY", "pneumonia")
best("NN", "pneumonia")
```


Submission 
```{r}
source("http://d396qusza40orc.cloudfront.net/rprog%2Fscripts%2Fsubmitscript3.R")
submit()

```


```{r}
?read.csv
```


```{r}
  df  <- read.csv('./outcome-of-care-measures.csv',
      colClasses=c('Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia'='character',
                   'Hospital.Name'='character'))
head(df$State)
```


```{r}
class(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack[1:10])
```


```{r}
class(df$State)
```


```{r}
"BB" %in% df$State ## To Check if State is valid.
"AL" %in% df$State
```


```{r}
head(df$Hospital.Name)
```


```{r}
outcomes   <- c("heart attack","heart failure","pneumonia")
"heart attack" %in% outcomes
"hert attack" %in% outcomes
```

```{r}
if (!("BB" %in% df$State)) {
  stop("invalid state")
}
```


```{r}
if (!("hert attack" %in% outcomes)) {
  stop("invalid outcome")
}
```

```{r}
df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack
str(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)
```

Find NA values
```{r}
df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == "Not Available"
```


```{r}
df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack[1:10]
#as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)[1:10]
#as.character(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)[1:10]
as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)[1:10]
```

Max and Min
[refer-1](http://stat.ethz.ch/R-manual/R-devel/library/base/html/Extremes.html)



```{r}
max(as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack))
min(as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack))
max(as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack),na.rm = T)
min(as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack),na.rm = T)

```


```{r}
max(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T)
min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T)
```


```{r}
df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T)
```

```{r}
any(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T))
```

```{r}
sum(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T))
```


Number of matches found
```{r}
sum(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T))
```


```{r}
df[df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T), 2]
```

Find the details for Texas
```{r}
df1  <- df[df$State == "TX",]
```


```{r}
df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T), 2]
```


Resolve Conflict by choosing the first hospital in alphabetical order
```{r}
min(df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack,na.rm = T), 2])

```




Find the details for MD
```{r}
df1  <- df[df$State == "MD",]
```


```{r}
df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia,na.rm = T), 2]
```

Find the details for NY
```{r}
df1  <- df[df$State == "NY",]
```


```{r}
df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia,na.rm = T), 2]
```

```{r}
min(df1[df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia,na.rm = T), 2])

min(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia)
sum(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == 10.0)
sum(df1$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia == "10.0")
```


========================================================
Qustion#3
##Ranking hospitals by outcome in a state


Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num). The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the ranking specified by the num argument. For example, the call

```{r}
rm(list = ls())
```


```{r}
rankhospital  <- function(state, outcome, num) {
  # Read the file outcome-of-care-measures.csv
  df  <- read.csv('./outcome-of-care-measures.csv',
      colClasses=c('Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia'='character',
                   'Hospital.Name'='character'))

  df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)

  df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure)

  df$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia)
  
  outcomes  <- c("heart attack", "heart failure", "pneumonia")
  
  if (!(state %in% df$State)) {
    stop("invalid state")
  }
  
  if (!(outcome %in% outcomes)) {
    stop("invalid outcome")
  }
  
  
  if(outcome == "heart attack") {
  df1  <- df[df$State == state,c(2,11)]
  names(df1)[2]  <- "Rate"
  }
  
  if(outcome == "heart failure") {
    df1  <- df[df$State == state,c(2,17)]
    names(df1)[2]  <- "Rate"
  }
  
  
  if(outcome == "pneumonia") {
    df1  <- df[df$State == state,c(2,23)]
    names(df1)[2]  <- "Rate"
  }
  # Remove rows where Ratings are not known.
  df1  <- df1[!is.na(df1$Rate),]
  # Order the dataframe by Names to help in Ranking
  df1  <- df1[order(df1[1]),]
  #head(df1)
  #tail(df1)
  
  # Set Rank 
  df1$rank  <- rank(df1$Rate,ties.method="first")
  
  # Return the values based on data provided
  if (num == "best") {
  return (df1[df1$rank == min(df1$rank),1])
  } else if (num == "worst") {
    return (df1[df1$rank == max(df1$rank),1])
  } else if ((num <= max(df1$rank)) | (num >= min(df1$rank))) {
    return (df1[df1$rank == num,1])
  } else {
    return (NA)
  }
  
}
```


The function reads the outcome-of-care-measures.csv file

```{r}
df  <- read.csv('./outcome-of-care-measures.csv',
      colClasses=c('Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure'='character',
                   'Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia'='character',
                   'Hospital.Name'='character'))

df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)

df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure)

df$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia  <- as.numeric(df$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia)

```

```{r}
class(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack)
class(df$Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure)
class(df$Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia)
```

```{r}
state  <- "NY"
outcome  <- "heart attack"
outcome  <- "heart failure"
outcome  <- "penumonia"
outcomes  <- c("heart attack", "heart failure", "pneumonia")
```

refer
[link1](http://stackoverflow.com/a/1296745/2356016)

```{r}

if(outcome == "heart attack") {
  df1  <- df[df$State == state,c(2,11)]
  names(df1)[2]  <- "Rate"
}

if(outcome == "heart failure") {
  df1  <- df[df$State == state,c(2,17)]
  names(df1)[2]  <- "Rate"
}


if(outcome == "pneumonia") {
  df1  <- df[df$State == state,c(2,23)]
  names(df1)[2]  <- "Rate"
}
# Remove rows where Ratings are not known.
df1  <- df1[!is.na(df1$Rate),]
# Order the dataframe by Names to help in Ranking
df1  <- df1[order(df1[1]),]
head(df1)
tail(df1)
```

```{r}
?rank
```

```{r}
?order
```


```{r}
class(df)
```

Assign rank to dataframe
```{r}
df1$rank  <- rank(df1$Rate,ties.method="first")
```


Test the results by viewing the results by sorting by ranks 
```{r}
df1[order(df1[3]),]
head(df1[order(df1[3]),],20)
tail(df1[order(df1[3]),],20)
```


Find min max and num specific ranking rows 
```{r}
df1[df1$rank == max(df1$rank),1]
df1[df1$rank == min(df1$rank),1]
df1[df1$rank == 1,1]
df1[df1$rank == 1.0,1]
df1[df1$rank == 171,1]
```


Return the results
```{r}
if (num == "best") {
  return (df1[df1$rank == min(df1$rank),1])
} else if (num == "worst") {
  return (df1[df1$rank == max(df1$rank),1])
} else if ((num <= max(df1$rank)) | (num >= min(df1$rank))) {
  return (df1[df1$rank == num,1])
} else {
  return (NA)
}
```

```{r}

num  <-  "best"

if (num == "best") {
  return(df1[df1$rank == min(df1$rank),1])
}
```


Remove results with NA ratings 
```{r}
df1[!is.na(df1$Rate),]
```


```{r}
str(df)
```

```{r}
names(df)
```

```{r}
head(df)
```

```{r}
grep("^Hospital.30.Day.Death..Mortality..Rates.from",names(df))
names(df)[grep("^Hospital.30.Day.Death..Mortality..Rates.from",names(df))]
```



