Title
========================================================


```{r}
library(data.table)
```

```{r}
?data.table
```

```{r}
df  <- data.frame(x = c("b","b","b","a","a"),v=rnorm(5))
df
```


```{r}
dt  <- data.table(x = c("b","b","b","a","a"),v=rnorm(5))
dt 
```

```{r}
CARS  <- data.table(cars)
head(CARS)
```

```{r}
tables()
```


```{r}
dt
```


```{r}
dt[2,]
```



```{r}
dt[dt$x == "b",]
```



```{r}
dt["b",]
```


```{r}
setkey(dt,x)
```

```{r}
dt
```


```{r}
haskey(dt)
```

```{r}
key(dt)
```



```{r}
dt["b",]
```

```{r}
dt["b",mult="first"]
```

```{r}
dt["b",mult="last"]
```


```{r}
dt["b"]
```


```{r}
grpsize  <- ceiling(1e7/26^2)
grpsize
```


```{r}
tt = system.time(df  <- data.frame(
  x = rep(LETTERS, each=26*grpsize),
  y = rep(letters, each=grpsize),
  v = runif(grpsize*26^2),
  stringsAsFactors=F
  ))
tt
```


```{r}
head(df,3)
```




```{r}
tail(df,3)
```



```{r}
dim(df)
```

Extract arbitrary group from df


```{r}
tt = system.time(
  ans1  <- df[df$x == "R" & df$y == "h",]
  )
tt
```

```{r}
head(ans1,3)
```

```{r}
dim(ans1)
```
Convert to data.table 

```{r}
dt  <- as.data.table(df)
system.time(setkey(dt,x,y))
```

Binary search
```{r}
ss = system.time(
  ans2  <- dt[J("R","h")]
  )
ss
#tt
```

vector scan

```{r}
system.time(
  ans1  <- dt[x == "R" & y == "h",]  
  )
```

```{r}
system.time(
  ans2  <- df[df$x == "R" & df$y == "h",]  
  )
```

```{r}
identical(
  dt[J("R","h"),],
  dt[data.table("R","h"),]
  )
```

Fast grouping 


```{r}
head(dt)
dt[,sum(v)]
```
When we supply a j expression and a ’by’ list of expressions, the j expression is repeated for each ’by’ group:


```{r}
dt[,sum(v),by=x]
```
The by in data.table is fast. Let’s compare it to tapply.


```{r}
ttt=system.time(tt <- tapply(dt$v,dt$x,sum)); ttt
```

```{r}
sss=system.time(ss <- dt[,sum(v),by=x]); sss
```

```{r}
head(tt)
```

```{r}
head(ss)
```

```{r}
as.vector(tt)
```

```{r}
ss$V1
```


```{r}
identical(as.vector(tt), ss$V1)
```

Next, let’s group by two columns:


```{r}
ttt=system.time(tt <- tapply(dt$v,list(dt$x,dt$y),sum)); ttt
```


```{r}
sss=system.time(ss <- dt[,sum(v),by="x,y"]); sss
```


```{r}
tt[1:5,1:5]
```

```{r}
head(ss)
```


```{r}
identical(
  as.vector(t(tt)),
  ss$V1
  )


```


t(tt)


==========================================================
http://stackoverflow.com/questions/13493124/outer-join-data-table-r

```{r}
a <- data.table(a=c(1,2,3),b=c(3,4,5)); a
b <- data.table(a=c(1,2),k=c(1,2)); b
setkey(a,a)
setkey(b,a)

key(a)
key(b)
```

```{r}
b[a,]
a[b,]
```


TODO:http://stackoverflow.com/questions/12773822/why-does-xy-join-of-data-tables-not-allow-a-full-outer-join-or-a-left-join







================================================

```{r}
require(data.table)

```


```{r}
# create a regular data.frame
theDF <- data.frame(A=1:10,
                     B=letters[1:10],
                     C=LETTERS[11:20],
                     D=rep(c("One", "Two", "Three"), length.out=10))
theDF
```

```{r}
# create a data.table
theDT <- data.table(A=1:10,
                     B=letters[1:10],
                     C=LETTERS[11:20],
                     D=rep(c("One", "Two", "Three"), length.out=10))
 # print them and compare
 theDF
```

Multiple columns in a data.frame should be specified as a character vector. With data.tables the columns should be specified as a list of the actual names, not as characters.

```{r}
theDT[, list(A, C)] 
```


```{r}
# just one column
theDT[, B]
```

one column while maintaining data.table structure
```{r}
theDT[, list(B)]
```

If we must specify the column names as characters (perhaps because they were passed as arguments to a function), the with argument should be set to FALSE.


```{r}
theDT[, "B", with = FALSE]
```

```{r}
theDT[, c("A", "C"), with = FALSE]
```

This time we used a vector to hold the column names instead of a list. These nuances are important to proper functions of data.tables but can lead to a great deal of frustration.

##Keys

show tables
```{r}
tables()
```
The key is used to index the data.table and will provide the extra speed.

set the key

```{r}
setkey(theDT, D)
```

show the data.table again

```{r}
theDT
```

The data have been reordered according to column D, which is sorted alphabetically.


```{r}
key(theDT)
```

```{r}
tables()
```


This adds some new functionality to selecting rows from data.tables. In addition to selecting rows by the row number or by some expression that evaluates to TRUE or FALSE, a value of the key column can be specified.



```{r}
theDT["One", ]
```


```{r}
theDT[c("One", "Two"), ]
```

More than one column can be set as the key.

```{r}
#install.packages("ggplot2")
require(ggplot2)
data(diamonds)
diamondsDT <- data.table(diamonds)
setkey(diamondsDT, cut, color)
```


```{r}
head(diamondsDT)
```


To access rows according to both keys, there is a special function named J. It takes multiple arguments, each of which is a vector of values to select.

```{r}
diamondsDT[J("Ideal", "E"), ]
```

```{r}
diamondsDT[J("Ideal", c("E", "D")), ]
```


##data.table Aggregation

The primary benefit of indexing is faster aggregation. While aggregate and the various d*ply functions will work because data.tables are just enhanced data.frames, they will be slower than using the built-in aggregation functionality of data.table.



we calculate the mean price of diamonds for each type of cut.
```{r}
aggregate(price ~ cut,  diamonds, mean)
```

To get the same result using data.table, we do this:


```{r}
diamondsDT[, mean(price), by = cut]
```
The only difference between this and the previous result is that the columns have different names. To specify the name of the resulting column, pass the aggregation function as a named list.


```{r}
diamondsDT[, list(price = mean(price)), by = cut]
```


To aggregate on multiple columns, specify them as a list().


```{r}
diamondsDT[, mean(price), by = list(cut, color)]
```

To aggregate multiple arguments, pass them as a list. Unlike with aggregate, a different metric can be measured for each column.


```{r}
diamondsDT[, list(price = mean(price), carat = mean(carat)), by = cut]
```




```{r}
diamondsDT[, list(price = mean(price), carat = mean(carat), caratSum = sum(carat)), by = cut]
```

Finally, both multiple metrics can be calculated and multiple grouping variables can be specified at the same time.


```{r}
diamondsDT[, list(price = mean(price), carat = mean(carat)), by = list(cut, color)]
```


